---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# lampcheck

`lampcheck` is an R package for quick data-quality checks and accessible,
Pixar-inspired visualization. It is designed to support exploratory data
analysis by helping users identify potential issues in a dataset before
analysis and by providing a consistent plotting theme.

The package was developed using Pixar film data, but its main functions can
be used with other datasets as well.

## Installation

You can install the development version of `lampcheck` from GitHub with:

``` r
# install.packages("devtools")
devtools::install_github("LukeRominiecki230/lampcheck")
```

## Data-quality checks

The `lamp_check()` function gives a quick overview of a data frame before
analysis. It reports the number of rows, columns, duplicated rows, and
incomplete rows, then returns a variable-level summary of missingness and
unique values.

```{r}
library(lampcheck)

lamp_check(pixar_films)
```

This kind of check can help identify potential problems before they affect
later summaries or visualizations.

## Pixar film data

`lampcheck` includes two datasets used throughout the package examples.

`pixar_films` contains basic information about Pixar films, including release
date, runtime, and MPA rating.

`public_response` contains several measures of public and critical reception,
including Rotten Tomatoes, Metacritic, CinemaScore, and Critics Choice scores.

```{r}
head(pixar_films)
```

The two datasets can be combined using the shared `film` variable.

```{r}
pixar_ratings <- dplyr::left_join(
  pixar_films,
  public_response,
  by = "film"
)

head(pixar_ratings)
```

## Lightyear plotting theme

`theme_lightyear()` is a custom `ggplot2` theme inspired by Buzz Lightyear.
The package's main brand colors are purple (`#5A2D82`) and yellow-green
(`#D7F171`).

```{r, fig.alt="Scatterplot showing Pixar film release dates and Rotten Tomatoes scores."}
ggplot2::ggplot(
  pixar_ratings,
  ggplot2::aes(x = release_date, y = rotten_tomatoes)
) +
  ggplot2::geom_point(
    size = 3,
    color = "#5A2D82",
    na.rm = TRUE
  ) +
  ggplot2::labs(
    title = "Pixar Films on Rotten Tomatoes",
    x = "Release date",
    y = "Rotten Tomatoes score"
  ) +
  theme_lightyear()
```

The plotting theme uses the package colors as accents while retaining a light
plotting background. This keeps the visual identity recognizable without
making the styling compete with the data.

## Branding and accessibility

The `lampcheck` brand is inspired by Buzz Lightyear's purple and bright
yellow-green color scheme. Purple (`#5A2D82`) is used as the primary
background and dark accent, while yellow-green (`#D7F171`) is used as the
primary foreground and bright accent.

The package separates its general brand styling from its statistical plotting
style. The `_brand.yml` file uses the purple and yellow-green pair strongly,
while `theme_lightyear()` keeps the main plotting panel white and uses the
brand colors more selectively. This preserves readability and leaves room for
color to represent data when needed.

The purple and yellow-green foreground/background pairing has a contrast
ratio of approximately 7.84:1, meeting the WCAG AAA threshold for normal
text. The brand also uses a sans-serif font family to keep the typography
simple and broadly readable.

## Vignette

For a fuller example of a Pixar analysis using `lamp_check()`,
`theme_lightyear()`, and common data moves, see:

``` r
vignette("my-vignette", package = "lampcheck")
```
