---
title: "Checking and Exploring Pixar Films with lampcheck"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Checking and Exploring Pixar Films with lampcheck}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

```{r setup, message=FALSE, warning=FALSE}
library(lampcheck)
library(dplyr)
library(ggplot2)

data("pixar_films", package = "lampcheck")
data("public_response", package = "lampcheck")
```

## Getting started

`lampcheck` provides simple tools for checking data before analysis and for
creating accessible, Pixar-inspired visualizations. This vignette uses Pixar
film and public-response data from the March 11, 2025 TidyTuesday dataset to
demonstrate a basic exploratory workflow.

The two datasets contain different kinds of information. `pixar_films`
contains release information such as runtime and MPA rating, while
`public_response` contains several measures of critical and audience
reception. Before combining or analyzing them, it is useful to inspect the
data for potential quality issues.

## Start with a data check

```{r lamp-check}
lamp_check(pixar_films)
```

The check shows that the film dataset contains no duplicated rows, but it does
contain some missing values. Rather than automatically removing every
incomplete observation, the next steps only remove records that cannot be
identified by film title.

## Filter records that cannot be identified

Because film title is the variable used to connect the two datasets, a missing
film title prevents a meaningful match.

```{r filter}
pixar_films_clean <- pixar_films |>
  filter(!is.na(film))
```

This keeps observations with missing values in other variables, such as
runtime, because those values may still be useful for analyses that do not
depend on runtime.

## Join film information with public response

The two datasets share the variable `film`, so a left join can add the
reception measures to the film information while retaining the Pixar film
records.

```{r join}
pixar_ratings <- pixar_films_clean |>
  left_join(public_response, by = "film")
```

A second data check shows what the combined analysis table looks like after
the join.

```{r joined-check}
lamp_check(pixar_ratings)
```

## Create variables for exploration

A release year is easier to use in summaries and plots than a complete date.
It is also useful to compare Rotten Tomatoes and Metacritic scores directly.

```{r mutate}
pixar_ratings <- pixar_ratings |>
  mutate(
    release_year = as.integer(format(release_date, "%Y")),
    critic_score_gap = rotten_tomatoes - metacritic
  )
```

Positive values of `critic_score_gap` indicate that a film's Rotten Tomatoes
score is higher than its Metacritic score, while negative values indicate the
opposite.

## Arrange films by critical reception

Sorting the joined data makes it easy to see which films received the highest
Rotten Tomatoes scores.

```{r arrange}
pixar_ratings |>
  arrange(desc(rotten_tomatoes)) |>
  select(film, release_year, rotten_tomatoes, metacritic) |>
  head(10)
```

This view is useful for identifying individual films that stand out before
moving to broader summaries.

## Summarize reception by film rating

Pixar films can also be compared by their MPA rating. Grouping and summarizing
the data provides a simple comparison of average Rotten Tomatoes scores across
rating categories.

```{r group-summarise}
rating_summary <- pixar_ratings |>
  group_by(film_rating) |>
  summarise(
    films = n(),
    mean_rotten_tomatoes = mean(rotten_tomatoes, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(desc(mean_rotten_tomatoes))

rating_summary
```

These averages are descriptive rather than causal. They provide one way to
look for patterns that could motivate additional exploration.

## Visualize Pixar's release history

The package's `theme_lightyear()` function uses the purple and yellow-green
palette inspired by Buzz Lightyear while keeping the plotting panel light and
readable.

```{r plot, fig.alt="Scatterplot of Pixar film release year against Rotten Tomatoes score, with films separated into panels by MPA film rating."}
ggplot(
  pixar_ratings,
  aes(x = release_year, y = rotten_tomatoes)
) +
  geom_point(
    size = 3,
    color = "#5A2D82",
    na.rm = TRUE
  ) +
  facet_wrap(~ film_rating) +
  labs(
    title = "Pixar Films on Rotten Tomatoes",
    subtitle = "Critical reception across Pixar's release history",
    x = "Release year",
    y = "Rotten Tomatoes score"
  ) +
  theme_lightyear()
```

The custom theme uses the package's Buzz Lightyear-inspired colors for visual
identity, but it does not force those colors onto the data itself. This keeps
the theme flexible for plots in which color may need to encode another
variable.

## Putting the workflow together

This exploration demonstrates a workflow in which data quality is considered
before analysis rather than after a problem appears. `lamp_check()` provides
a quick view of missingness and other basic characteristics, while the data
moves build an analysis table, create useful variables, identify notable
observations, and summarize broader patterns. `theme_lightyear()` then provides
a consistent visual style for communicating the results.
